---
title: "Consumer and Firm Behavior, Assignment1"
author: "CSORNAI, Gyula"
date: "February 18, 2017"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## TASK9: Boris's morining coffee

### Short summary

Blue line shows the maximum amount of coffe and milk combinations Boris can chose. 
At the maximum he can buy 18oz milk and 9oz coffe on 1$ coffee price: *MILK = 18-2COFFEE*.
However he prefers 4 part coffe, 1 part milk ( *MILK=1/4COFFEE*), so on this price his optimal quantities are

*1/4COFFE = 18-2 x COFFEE -> 2.25COFFEE = 18 ->* **COFFEE = 8, MILK=2**

If coffe prices goes from 1$ to 3.25 USD, his budget line changes:

*MILK=18-18/(9/3.25)COFFEE -> MILK=18-6.75COFFEE* 

so

*1/4COFFEE = 18-6.5COFFE -> 6.75COFFEE = 18 ->* **COFFEE=2.67, MILK=0.67**


#### Visualization


```{r pressure, echo=FALSE}
library(ggplot2)
icp2 <- (-1)*18/(9/3.25)
ggplot()+
  xlim(0,11)+
  ylim(-4,20)+
  labs(
    x='Price of coffee',
    y='Price of milk',
    title='Consumer choices of Boris related to coffee and milk'
  )+
  geom_hline(aes(yintercept=0),col='black')+
  geom_vline(aes(xintercept=0),col='black')+
  geom_abline(intercept = 18, slope = -2, col='dodgerblue3')+
  geom_abline(intercept = 18, slope = icp2, col='indianred')+
  geom_vline(aes(xintercept=8),col='chocolate4',linetype='dashed')+
  geom_hline(aes(yintercept=2),col='chocolate4',linetype='dashed')+
  geom_vline(aes(xintercept=2.67),col='palevioletred4',linetype='dotted')+
  geom_hline(aes(yintercept=0.67),col='palevioletred4',linetype='dotted')+
  geom_label(aes(x=0, y=18, label='Max milk volume\n for budget=18oz'),nudge_x = 1, nudge_y=1, size=3)+
  geom_label(aes(x=9, y=0, label='Max coffe volume\n for budget on\n 1$ price = 9oz'),nudge_x = 0, nudge_y=-2, size=3)+
  geom_label(aes(x=3, y=0, label='Max coffe volume\n for budget on\n 3.25$ price = 9oz'),nudge_x = 0, nudge_y=-2, size=3)+
  geom_label(aes(x=8, y=2, label='Preference point for budget\non 1$ coffee price: 8oz coffee,\n 2oz milk'),nudge_x = 1.5, nudge_y=2, size=3)+
  geom_label(aes(x=2.67, y=0.67, label='Preference point for budget\non 3.25$ coffee price:\n 2.67oz coffee, 0.67 oz milk '),nudge_x = 2, nudge_y=3, size=3)+
  theme_bw()
```

## TASK10: Sam and the oranges - confused...

### Optimal consumption point, original prices

* Sam's max budget: *6$*
* The maximum cup OJ Sam can buy on 2$/cup price: *3*
* The maximum cup AJ Sam can buy on 1$/cup price: *6*
* Equation of budget line: *OJ = 3-0.5AJ*
* Substitution: *1OJ = 3 AJ*
* Therefore optimal point would be: *3-0.5AJ = 3 AJ -> 3.5AJ = 3 ->* **AJ=`r 3/3.5 ` , OJ=`r 9/3.5` **
* But if cup is a unit: **AJ=1, OJ=2** , which is under the budget line (*1$ + 2x2$ = 5$*)

### Optimal consumption point, AJ=2$ prices

* Sam's original bundle, full cups: **AJ=1, OJ=2**
* This costed *5$* on *original price*
* This costs *6$* on *new price* (so most possibly this is not the question)
* Sam's original bundle, part cups: **AJ=0.85 , OJ=2.57**
* After price change this would cost him *0.85x2 + 2.57x2 =* **6.84**
* The new budget line is * OJ =3-AJ * 
* The new optimal point would be *3AJ = 3-AJ -> 4AJ = 3 ->* *AJ=0.75, OJ=2.25*
* The new optimal point would cost him *2 x (0.75 + 2.25) = * *6$* 
* This means **(?)** that we need to add him **0.84$** if he wants to afford original consumption, given not full cups, and we do not need to add him anything if he can only drink full cups.

## TASK11: Bruce, bread and butter

### Original optimal choice

* Total budget: **6$**
* If bagel price is 2$ and butter is 1$, the budget line is *BAGEL = 3 - 0.5BUTTER*
* Bruce's preference line is *BAGEL=BUTTER*
* Bruce's optimal bundle is *BUTTER = 3-0.5BUTTER -> 1.5BUTTER = 3 -> * **BUTTER=2, BAGEL=2**

### Income change needed

* If bagel price is 2$ and butter is also 2$ , having 6$ in hand, the budget line is *BAGEL = 3 - BUTTER*
* Bruce's riginal optimal bundle is *BUTTER=2, BAGEL=2*, this would cost him now *8$*
* Therefore Bruce would need **2$** to afford the new bundle
* If he had this 8$, his budget line would be *BAGEL = 4 - BUTTER*, so with this package the bundle is again **BUTTER=2, BAGEL=2**

### Substitution and income effect
* On this price, Bruce could afford *BAGEL=1.5, BUTTER=1.5* from 6$
* **Substitution effect**: *BUTTER(original)-BUTTER(original package, increased price) = * **0 package/week**
* **Income effect**: *BUTTER(original package, increased price)-BUTTER(original budget, increased price) = * **0.5package/week**

## TASK12: Golden Gate toll

### A: Revenue

* If *Q=10000-1000P* and toll fee = *3$*
* The revenue is *R=PxQ* so (10000-3x1000)x3 = ** 21,000 **

### B: Price elasticity

$$\epsilon = \frac{\frac{P1-P}{P}}{\frac{Q1-Q}{Q}}$$


*  Price elasticity of demand is defined as the percentage change in the quantity of a good demanded that results from a 1 percent change in price.
* Q = 7000, P = 3, P1 = 1.01 * 3 = 3.03, Q1 = 10000-3.03x1000 = 6970
* Elasticity = ( 0.01/ (-130/7000) ) = **-0.54**

### C: Revenue change

* Could authorities increase their price?
* **YES**, because price is relatively inelastic, **R1** = P1xQ1 = 6970x3.03 = **21,119.1** > 21,000
* This is because there is no substitution product.

### D: Impact of Red and White ferry

* The elasticity of the demand would increase
* In general, the absolute value of price elasticity will rise with the availability of attractive substitutes.

## TASK13: Elasticity of a curve

### Elasticity calculations

#### Original equation 

$$Q = 100 - 0.5P$$
#### Demand curve 

$$P = 200 - 2Q$$

#### Calculate elasticity

$$\epsilon = \frac{P}{Q} * \frac{1}{slope} = \frac{alfa+beta*Q}{Q} * \frac{1}{beta}$$

#### Unit elasticity point 
At inique elasticity
$$-1 = \frac{alfa+beta*Q}{Q} * \frac{1}{beta}$$
$$Q= -\frac{alfa}{ 2*beta} $$

#### Demand curve chart

```{r elasticityCheck, echo=FALSE}
alfa<-200
beta<- -2
qfp <- function(q){return(alfa+beta*q)}
qelasticity <- (-1*alfa) / (2*beta)


labl <- paste(paste(
  paste('Unit elasticity point: [',as.character(qelasticity),sep= ""),
  as.character(qfp(qelasticity)),sep=","),"]",sep=" ")
ggplot()+
  xlim(0,110)+
  ylim(-2,210)+
  labs(
    x='Q(antity)',
    y='P(rice)',
    title='Demand curve for the qiven equation'
  )+
  geom_hline(aes(yintercept=0),col='black')+
  geom_vline(aes(xintercept=0),col='black')+
  geom_abline(intercept = alfa, slope = beta, col='dodgerblue3')+
  geom_vline(aes(xintercept=qelasticity),col='palevioletred4',linetype='dotted')+
  geom_hline(aes(yintercept=qfp(qelasticity)),col='palevioletred4',linetype='dotted')+
  geom_label(aes(x=qelasticity, y=qfp(qelasticity)), label=labl ,nudge_x = 25, nudge_y=15, size=5)+
  geom_label(aes(x=(qelasticity - 30) , y=(qfp(qelasticity)+30)), label='Elastic' ,nudge_x = 10, nudge_y=30, size=5)+
  geom_label(aes(x=(qelasticity + 30) , y= (qfp(qelasticity) - 30)), label='Not elastic' ,nudge_x =15, nudge_y=-30, size=5)+
  theme_bw()


```

### Maximize expanditure

The maximum expenditure is expected in the unit elasticity point, because that is the point, where 1% price increase will result 1% quantity loss, which means that the gain is approximately the same as the loss.
## TASK14: Flights to Budapest

* At price 40$ quantity is 300
* At price 60$ quantity is 280

### Demand equation

```{r demandeq, echo=FALSE}
p1<- 40
p2<- 60
q1<-300
q2<-280
beta <- (p2-p1) / (q2-q1)
alfa <- (-1)*q2*beta+p2
e1 <- (p1/q1) * (1/beta) 
e2 <- (p2/q2) * (1/beta)
lab1 <- paste(
  paste("(","):\n", sep=paste(as.character(q1),as.character(p1),sep=",")),
  paste("E1: ",substring(as.character(e1),1,6))
)
lab2 <- paste(
  paste("(","):\n", sep=paste(as.character(q2),as.character(p2),sep=",")),
  paste("E2: ",substring(as.character(e2),1,6))
)
##How to use equation dynmically?!
```
$$P=340-1*Q$$


```{r demandplot}
ggplot()+
  xlim(0, (-1*(alfa/beta) + 30))+
  ylim(-2,alfa+10)+
  labs(
    x='Q(antity)',
    y='P(rice)',
    title='Demand curve for Budapest flights'
  )+
  geom_hline(aes(yintercept=0),col='black')+
  geom_vline(aes(xintercept=0),col='black')+
  geom_abline(intercept = alfa, slope = beta, col='dodgerblue3')+
  geom_vline(aes(xintercept=q1),col='palevioletred4',linetype='dotted')+
  geom_hline(aes(yintercept=p1),col='palevioletred4',linetype='dotted')+
  geom_vline(aes(xintercept=q2),col='red4',linetype='dotted')+
  geom_hline(aes(yintercept=p2),col='red4',linetype='dotted')+
  geom_label(aes(x=(q2 - 20) , y=p2+20), label=lab2 ,nudge_x = -10, nudge_y=10, size=3)+
  geom_label(aes(x=q1 + 20  , y= p2 - 20), label=lab1 ,nudge_x = 10, nudge_y=10, size=3)+
  theme_bw()

```
