---
title: "Consumer and Firm Behavior, Assignment1"
author: "CSORNAI, Gyula"
date: "February 21, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```
## TASK1: Opportunity cost of lending money

* Amount of money lent: *500$*
* Returned money+goods: R=*505$*
* Amount of money I had on the savings account with 6% annual interest rate: P=`r 500*1.06`$
* Opportunity cost (P-R)= **`r 500*1.06 - 505`$**

## TASK2: Dana and the rock concert

* The concert ticket is a **sunk cost**, because at the point of the decision it is already purchased.
* Before buying the ticket the marginal cost of the party is 0$
* After buying the ticket the marginal cost of the party is still 0$ because additional parties will not cost this much.
* If she is rational, she will go to the welcome party, because sunk costs should not be considered in present decisions, only the costs and benefits observed in the present and future. Going to the concert is not a future cost.
* If she could resell the ticket at any price, her benefit of going to the party would be:
  + B~party_preference~ > B~concert_preference~ without the ticket, because of preference
  + B~party~' = B~party_preference~ + p~resell~ > B~comcert_preference~ = B~concert~
  
## TASK3: Adding RAM to computer

### With 100$/1GB RAM

|Number of added RAM GB| Total benefit ($) | Average benefit per RAM ($/GB)| Marginal benefit ($/GB) | TOTAL |
|:--------------------:|:-----------------:|:-----------------------------:|:-----------------------:|:-----:|
| 0                    |0                  |0                              | 400                     | +300  |
| 1                    | 400               | 400                           |                         |       |
|                      |                   |                               | 200                     | +100  |
| 2                    | 600               | 300                           |                         |       |
|                      |                   |                               | 100                     | 0     |
| 3                    | 700               | 233.33                        |                         |       |
|                      |                   |                               |  50                     | -50   |
| 4                    | 750               | 187.5                         |                         |       |

Based on the calculation we should purchase 3G memory, but not more.

### With 300$/1GB RAM

In this case we should only buy 1GB only, because in the second case the marginal benefit is smaller than marginal cost.


## TASK4: DVD supply

### Supply curve

$$ P= 2Q$$

### Demand curve

$$ P=42-Q $$

### Units traded on different prices

```{r supdim, echo=FALSE}
alfa1<-0
alfa2<-42
beta1<-2
beta2<- -1
c <- 42

p1<- 35
p2 <- 14

suppQty <- function(price){return(as.numeric((price-alfa1)/beta1))}
demdQty <- function(price){return(as.numeric((price-alfa2)/beta2))}
eqlbPrc <- function(){return((alfa1*beta2-alfa2*beta1)/(beta2-beta1))}

suppLabl <- function(price){
  return(
    paste(
      paste(
        "The supply on price = ",
        "\nis: ",
        sep=as.character(price)),
      as.character(suppQty(price))
    )
  )
}

demdLabl <- function(price){
  return(
    paste(
      paste(
        "The demand on price = ",
        "\nis: ",
        sep=as.character(price)),
      as.character(demdQty(price))
    )
  )
}


ggplot()+
  xlim(0,max(alfa1,alfa2)*1.1)+
  ylim(-2,max(abs(alfa1/beta1),abs(alfa2/beta2))*1.1)+
  labs(
    x='Q(antity)',
    y='P(rice)',
    title='Demand curve for the qiven equation'
  )+
  geom_hline(aes(yintercept=0),col='black')+
  geom_vline(aes(xintercept=0),col='black')+
  geom_abline(intercept = alfa1, slope = beta1, col='dodgerblue3')+
  geom_abline(intercept = alfa2, slope = beta2, col='magenta4')+
  geom_label(aes(x=0, y=alfa2), label='Demand curve' ,nudge_x = 3, nudge_y=3, size=3)+
  geom_label(aes(x=0 , y=alfa1), label='Supply curve' ,nudge_x = 1, nudge_y=-2, size=3)+
  geom_hline(aes(yintercept=p1),col='palevioletred4',linetype='dotted')+
  geom_vline(aes(xintercept=suppQty(p1)),col='palevioletred4',linetype='dotted')+
  geom_vline(aes(xintercept=demdQty(p1)),col='palevioletred4',linetype='dotted')+
  geom_label(aes(x=demdQty(p1), y=p1), label=demdLabl(p1) ,nudge_x = 5, nudge_y=5, size=3)+
  geom_label(aes(x=suppQty(p1), y=p1), label=suppLabl(p1) ,nudge_x = 10, nudge_y=5, size=3)+
  geom_hline(aes(yintercept=p2),col='indianred',linetype='dotted')+
  geom_vline(aes(xintercept=suppQty(p2)),col='indianred',linetype='dotted')+
  geom_vline(aes(xintercept=demdQty(p2)),col='indianred',linetype='dotted')+
  geom_label(aes(x=demdQty(p2), y=p2), label=demdLabl(p2) ,nudge_x = 5, nudge_y=5, size=3)+
  geom_label(aes(x=suppQty(p2), y=p2), label=suppLabl(p2) ,nudge_x = 5, nudge_y=-5, size=3)+
  geom_hline(aes(yintercept=eqlbPrc()),col='seagreen4',linetype='dotted')+
  geom_vline(aes(xintercept=demdQty(eqlbPrc())),col='seagreen4',linetype='dotted')+
  geom_label(aes(x=demdQty(eqlbPrc()), y=eqlbPrc()), label=demdLabl(eqlbPrc()) ,nudge_x = 5, nudge_y=5, size=3)+
  theme_bw()

```


As shown on the graph on price `r as.character(p1)` the supply is **`r suppQty(p1)`** whereas the demand is **`r demdQty(p1)`**, so in this case **`r ifelse(suppQty(p1) > demdQty(p1),"sellers","buyers")`** are unhappy.

From the other side on price `r as.character(p1)` the supply is **`r suppQty(p1)`** whereas the demand is **`r demdQty(p1)`**, so in this case **`r ifelse(suppQty(p1) > demdQty(p1),"sellers","buyers")`** are unhappy.

### Equilibrium point

The equilkibrium point is when 
$$Q~demand~=Q~supply~$$
so

$$42-P=\frac{P}{2}$$

which results an equilibrium price =  **`r (alfa1*beta2-alfa2*beta1)/(beta2-beta1) `** 

### Total revenue

The total revenue is at the maximum at equilibrium price, so R = **`r eqlbPrc() * demdQty(eqlbPrc())`**


## TASK5: Price and quantity diagrams


```{r coil, echo=FALSE, fig.width=4, fig.height=3,out.extra='style="float:left"'}
ggplot()+
  xlim(-0.1,5.1)+
  ylim(-0.1,5.1)+
  labs(
    x='Q(antity)',
    y='P(rice)',
    title='Supply curve shift in case petroleum reserve changes',
    caption='As petroleum reserves decrease, 
it becomes more difficult to find and recover crude oil.
    Price goes up, demand decreases.'
  )+
  geom_hline(aes(yintercept=0),col='black')+
  geom_vline(aes(xintercept=0),col='black')+
  geom_abline(intercept = 5, slope = -1, col='dodgerblue3')+
  geom_abline(intercept = 0, slope = 1, col='indianred')+
  geom_label(aes(x=0, y=5), label='Demand curve' ,nudge_x = 1, nudge_y=-1, size=3)+
  geom_label(aes(x=0 , y=0), label='Supply curve' ,nudge_x = 1, nudge_y=1, size=3)+
  geom_abline(intercept = 1, slope = 1, col='red')+
  geom_label(aes(x=1 , y=0), label='NEW Supply curve' ,nudge_x =0, nudge_y=2, size=3)+
  theme_bw()
```


```{r airt, echo=FALSE, fig.width=4, fig.height=3,out.extra='style="float:left"'}
ggplot()+
  xlim(-0.1,5.1)+
  ylim(-0.1,5.1)+
  labs(
    x='Q(antity)',
    y='P(rice)',
    title='Demand shift due to worries around air travel',
    caption='Air travel: Worries about air 
safety cause travelers to shy away from air travel..
    Demand goes down, price decreases.'
  )+
  geom_hline(aes(yintercept=0),col='black')+
  geom_vline(aes(xintercept=0),col='black')+
  geom_abline(intercept = 5, slope = -1, col='dodgerblue3')+
  geom_abline(intercept = 0, slope = 1, col='indianred')+
  geom_label(aes(x=0, y=5), label='Demand curve' ,nudge_x = 1, nudge_y=-1, size=3)+
  geom_label(aes(x=0 , y=0), label='Supply curve' ,nudge_x = 1, nudge_y=1, size=3)+
  geom_abline(intercept = 4, slope = -1, col='blue')+
  geom_label(aes(x=0 , y=4), label='NEW Demand curve' ,nudge_x =1, nudge_y=-1, size=3)+
  theme_bw()
```


```{r railt, echo=FALSE, fig.width=4, fig.height=3, out.extra='style="float:left"'}
ggplot()+
  xlim(-0.1,6.1)+
  ylim(-0.1,6.1)+
  labs(
    x='Q(antity)',
    y='P(rice)',
    title='Demand shift due to worries around air travel 
    - in railroad travel',
    caption='Rail travel: Worries about air safety 
cause travelers to shy away from air travel
    Demand goes up, price increases'
  )+
  geom_hline(aes(yintercept=0),col='black')+
  geom_vline(aes(xintercept=0),col='black')+
  geom_abline(intercept = 5, slope = -1, col='dodgerblue3')+
  geom_abline(intercept = 0, slope = 1, col='indianred')+
  geom_label(aes(x=0, y=5), label='Demand curve' ,nudge_x = 1, nudge_y=-1, size=3)+
  geom_label(aes(x=0 , y=0), label='Supply curve' ,nudge_x = 1, nudge_y=1, size=3)+
  geom_abline(intercept = 6, slope = -1, col='blue')+
  geom_label(aes(x=0 , y=6), label='NEW Demand curve' ,nudge_x =1, nudge_y=-1, size=3)+
  theme_bw()
```

```{r hotel, echo=FALSE, fig.width=4, fig.height=3,out.extra='style="float:left"'}
ggplot()+
  xlim(-0.1,5.1)+
  ylim(-0.1,5.1)+
  labs(
    x='Q(antity)',
    y='P(rice)',
    title='Demand shift in hotel rooms in Hawaii due to worries around air travel',
    caption='Hotel rooms in Hawaii: 
Worries about air safety cause travelers to shy away from air travel.
    Demand goes down, price decreases.'
  )+
  geom_hline(aes(yintercept=0),col='black')+
  geom_vline(aes(xintercept=0),col='black')+
  geom_abline(intercept = 5, slope = -1, col='dodgerblue3')+
  geom_abline(intercept = 0, slope = 1, col='indianred')+
  geom_label(aes(x=0, y=5), label='Demand curve' ,nudge_x = 1, nudge_y=-1, size=3)+
  geom_label(aes(x=0 , y=0), label='Supply curve' ,nudge_x = 1, nudge_y=1, size=3)+
  geom_abline(intercept = 4, slope = -1, col='blue')+
  geom_label(aes(x=0 , y=4), label='NEW Demand curve' ,nudge_x =1, nudge_y=-1, size=3)+
  theme_bw()
```


```{r milk, echo=FALSE, fig.width=4, fig.height=3}
ggplot()+
  xlim(-0.1,5.1)+
  ylim(-1.1,5.1)+
  labs(
    x='Q(antity)',
    y='P(rice)',
    title='Supply shift in in milk price due to productivity',
    caption='A genetically engineered hormone enables 
large milk producers to cut production costs. 
    Prices go down, demand is more quantity 
(NOTE: only in case GMO milk is a substitution of bio milk).'
  )+
  geom_hline(aes(yintercept=0),col='black')+
  geom_vline(aes(xintercept=0),col='black')+
  geom_abline(intercept = 5, slope = -1, col='dodgerblue3')+
  geom_abline(intercept = 0, slope = 1, col='indianred')+
  geom_label(aes(x=0, y=5), label='Demand curve' ,nudge_x = 1, nudge_y=-1, size=3)+
  geom_label(aes(x=0 , y=0), label='Supply curve' ,nudge_x = 0.8, nudge_y=1, size=3)+
  geom_abline(intercept = 0, slope = 0.8, col='red')+
  geom_label(aes(x=0 , y=-1), label='NEW Supply curve' ,nudge_x =1, nudge_y=1, size=3)+
  theme_bw()
```


## TASK6: Handball game

### Equilibrium price

#### Demand

$$ P=1900-\frac{1}{50}*Q$$

#### Supply

$$Q=19000$$

So price is 

$$P=1900-\frac{19000}{50}=1520 $$

### Excess demand

On 50 USD price the quantity demanded would be

$$Q=(1900-50)*50 - 19000 = 73500$$

### Demand jump

If demand jumps the new equation:

$$Q=(2100-50)*50 - 19000 = 73500$$

### Price ceiling

```{r ceil, echo=FALSE}
ggplot()+
  xlim(-0.1,19200)+
  ylim(-0.1,1910)+
  labs(
    x='Q(antity)',
    y='P(rice)',
    title='Regulated vs nonregulated price',
    caption='As petroleum reserves decrease, 
it becomes more difficult to find and recover crude oil.
    Price goes up, demand decreases.'
  )+
  geom_hline(aes(yintercept=0),col='black')+
  geom_vline(aes(xintercept=0),col='black')+
  geom_abline(intercept = 1900, slope = -1/50, col='dodgerblue3')+
  geom_vline(aes(xintercept=19000),col='indianred')+
  geom_label(aes(x=1000, y=1800), label='Demand curve' ,nudge_x = 1, nudge_y=-1, size=3)+
  geom_label(aes(x=18000 , y=1000), label='Maximum quantity\nFixed supply' ,nudge_x =0, nudge_y=-0, size=3)+
  geom_abline(intercept = 0, slope = 0.1, col='red')+
  geom_hline(aes(yintercept=50),col='blue')+
  geom_label(aes(x=0 , y=0), label='Typical Supply curve' ,nudge_x =1500, nudge_y=100, size=3)+
  geom_label(aes(x=0 , y=50), label='Fixed price' ,nudge_x =10000, nudge_y=100, size=3)+
  theme_bw()
```



## TASK7: Marginal price changes

```{r margprc, echo=FALSE}
ggplot()+
  xlim(-0.1,4.1)+
  ylim(-0.1,10.1)+
  labs(
    x='Q(antity) - (lb/day)',
    y='P(rice) - ($/lb)',
    title='Supply curve changes tue to marginal cost increase',
    caption='The marginal cost doubled.'
  )+
  geom_hline(aes(yintercept=0),col='black')+
  geom_vline(aes(xintercept=0),col='black')+
  geom_abline(intercept = 8, slope = -2, col='dodgerblue3')+
  geom_abline(intercept = 0, slope = 2, col='indianred')+
  geom_label(aes(x=0, y=8), label='Demand curve' ,nudge_x = 1, nudge_y=-1, size=3)+
  geom_label(aes(x=0 , y=0), label='Supply curve' ,nudge_x = 1, nudge_y=1, size=3)+
  geom_abline(intercept = 0, slope = 4, col='red')+
  geom_label(aes(x=2 , y=8), label='NEW Supply curve' ,nudge_x =0.5, nudge_y=0.5, size=3)+
  theme_bw()
```


## TASK8: Martha's budget

### Budget constraint for Martha

```{r mathbudg, echo=FALSE}
ggplot()+
  xlim(-0.1,65)+
  ylim(-0.1,155)+
  labs(
    x='Malted milk ball quantity - 2.5$/unit',
    y='Composite good quantity - 1$/unit',
    title='Budget constraint for Martha'
  )+
  geom_hline(aes(yintercept=0),col='black')+
  geom_vline(aes(xintercept=0),col='black')+
  geom_abline(intercept = 150, slope = -2.5, col='dodgerblue3')+
  geom_label(aes(x=0, y=150), label='Martha\'s budget line' ,nudge_x = 10, nudge_y=-10, size=3)+
  theme_bw()
```

### Opportunity cost

* Opportunity cost for 1 more unit of composite goods is the number of Milk ball she ascrifices: **`r 1/2.5`**
* Opportunity cost for 1 more unit of milk ball is the number of composite good she ascrifices: **2.5**


## TASK9: Boris's morining coffee

### Short summary

Blue line shows the maximum amount of coffe and milk combinations Boris can chose. 
At the maximum he can buy 18oz milk and 9oz coffe on 1$ coffee price: *MILK = 18-2COFFEE*.
However he prefers 4 part coffe, 1 part milk ( *MILK=1/4COFFEE*), so on this price his optimal quantities are

*1/4COFFE = 18-2 x COFFEE -> 2.25COFFEE = 18 ->* **COFFEE = 8, MILK=2**

If coffe prices goes from 1$ to 3.25 USD, his budget line changes:

*MILK=18-18/(9/3.25)COFFEE -> MILK=18-6.75COFFEE* 

so

*1/4COFFEE = 18-6.5COFFE -> 6.75COFFEE = 18 ->* **COFFEE=2.67, MILK=0.67**


#### Visualization


```{r pressure, echo=FALSE}

icp2 <- (-1)*18/(9/3.25)
ggplot()+
  xlim(0,11)+
  ylim(-4,20)+
  labs(
    x='Price of coffee',
    y='Price of milk',
    title='Consumer choices of Boris related to coffee and milk'
  )+
  geom_hline(aes(yintercept=0),col='black')+
  geom_vline(aes(xintercept=0),col='black')+
  geom_abline(intercept = 18, slope = -2, col='dodgerblue3')+
  geom_abline(intercept = 18, slope = icp2, col='indianred')+
  geom_vline(aes(xintercept=8),col='chocolate4',linetype='dashed')+
  geom_hline(aes(yintercept=2),col='chocolate4',linetype='dashed')+
  geom_vline(aes(xintercept=2.67),col='palevioletred4',linetype='dotted')+
  geom_hline(aes(yintercept=0.67),col='palevioletred4',linetype='dotted')+
  geom_label(aes(x=0, y=18, label='Max milk volume\n for budget=18oz'),nudge_x = 1, nudge_y=1, size=3)+
  geom_label(aes(x=9, y=0, label='Max coffe volume\n for budget on\n 1$ price = 9oz'),nudge_x = 0, nudge_y=-2, size=3)+
  geom_label(aes(x=3, y=0, label='Max coffe volume\n for budget on\n 3.25$ price = 9oz'),nudge_x = 0, nudge_y=-2, size=3)+
  geom_label(aes(x=8, y=2, label='Preference point for budget\non 1$ coffee price: 8oz coffee,\n 2oz milk'),nudge_x = 1.5, nudge_y=2, size=3)+
  geom_label(aes(x=2.67, y=0.67, label='Preference point for budget\non 3.25$ coffee price:\n 2.67oz coffee, 0.67 oz milk '),nudge_x = 2, nudge_y=3, size=3)+
  theme_bw()
```

## TASK10: Sam and the oranges 

### Optimal consumption point, original prices

* Sam's max budget: *6$*
* The maximum cup OJ Sam can buy on 2$/cup price: *3*
* The maximum cup AJ Sam can buy on 1$/cup price: *6*
* Equation of budget line: *OJ = 3-0.5AJ*
* Substitution: *1OJ = 3 AJ*
* Therefore optimal point would be: *3-0.5AJ = 3 AJ -> 3.5AJ = 3 ->* **AJ=`r 3/3.5 ` , OJ=`r 9/3.5` **
* But if cup is a unit: **AJ=1, OJ=2** , which is under the budget line (*1$ + 2x2$ = 5$*)

### Optimal consumption point, AJ=2$ prices

* Sam's original bundle, full cups: **AJ=1, OJ=2**
* This costed *5$* on *original price*
* This costs *6$* on *new price* (so most possibly this is not the question)
* Sam's original bundle, part cups: **AJ=0.85 , OJ=2.57**
* After price change this would cost him *0.85x2 + 2.57x2 =* **6.84**
* The new budget line is * OJ =3-AJ * 
* The new optimal point would be *3AJ = 3-AJ -> 4AJ = 3 ->* *AJ=0.75, OJ=2.25*
* The new optimal point would cost him *2 x (0.75 + 2.25) = * *6$* 
* This means **(?)** that we need to add him **0.84$** if he wants to afford original consumption, given not full cups, and we do not need to add him anything if he can only drink full cups.

## TASK11: Bruce, bread and butter

### Original optimal choice

* Total budget: **6$**
* If bagel price is 2$ and butter is 1$, the budget line is *BAGEL = 3 - 0.5BUTTER*
* Bruce's preference line is *BAGEL=BUTTER*
* Bruce's optimal bundle is *BUTTER = 3-0.5BUTTER -> 1.5BUTTER = 3 -> * **BUTTER=2, BAGEL=2**

### Income change needed

* If bagel price is 2$ and butter is also 2$ , having 6$ in hand, the budget line is *BAGEL = 3 - BUTTER*
* Bruce's riginal optimal bundle is *BUTTER=2, BAGEL=2*, this would cost him now *8$*
* Therefore Bruce would need **2$** to afford the new bundle
* If he had this 8$, his budget line would be *BAGEL = 4 - BUTTER*, so with this package the bundle is again **BUTTER=2, BAGEL=2**

### Substitution and income effect
* On this price, Bruce could afford *BAGEL=1.5, BUTTER=1.5* from 6$
* **Substitution effect**: *BUTTER(original)-BUTTER(original package, increased price) = * **0 package/week**
* **Income effect**: *BUTTER(original package, increased price)-BUTTER(original budget, increased price) = * **0.5package/week**

## TASK12: Golden Gate toll

### A: Revenue

* If *Q=10000-1000P* and toll fee = *3$*
* The revenue is *R=PxQ* so (10000-3x1000)x3 = ** 21,000 **

### B: Price elasticity

$$\epsilon = \frac{\frac{P1-P}{P}}{\frac{Q1-Q}{Q}}$$


*  Price elasticity of demand is defined as the percentage change in the quantity of a good demanded that results from a 1 percent change in price.
* Q = 7000, P = 3, P1 = 1.01 * 3 = 3.03, Q1 = 10000-3.03x1000 = 6970
* Elasticity = ( 0.01/ (-130/7000) ) = **-0.54**

### C: Revenue change

* Could authorities increase their price?
* **YES**, because price is relatively inelastic, **R1** = P1xQ1 = 6970x3.03 = **21,119.1** > 21,000
* This is because there is no substitution product.

### D: Impact of Red and White ferry

* The elasticity of the demand would increase
* In general, the absolute value of price elasticity will rise with the availability of attractive substitutes.

## TASK13: Elasticity of a curve

### Elasticity calculations

#### Original equation 

$$Q = 100 - 0.5P$$

#### Demand curve 

$$P = 200 - 2Q$$

#### Calculate elasticity

$$\epsilon = \frac{P}{Q} * \frac{1}{slope} = \frac{alfa+beta*Q}{Q} * \frac{1}{beta}$$

#### Unit elasticity point 

At inique elasticity
$$-1 = \frac{alfa+beta*Q}{Q} * \frac{1}{beta}$$
$$Q= -\frac{alfa}{ 2*beta} $$


#### Demand curve chart

```{r elasticityCheck, echo=FALSE}
alfa<-200
beta<- -2
qfp <- function(q){return(alfa+beta*q)}
qelasticity <- (-1*alfa) / (2*beta)


labl <- paste(paste(
  paste('Unit elasticity point: [',as.character(qelasticity),sep= ""),
  as.character(qfp(qelasticity)),sep=","),"]",sep=" ")
ggplot()+
  xlim(0,110)+
  ylim(-2,210)+
  labs(
    x='Q(antity)',
    y='P(rice)',
    title='Demand curve for the qiven equation'
  )+
  geom_hline(aes(yintercept=0),col='black')+
  geom_vline(aes(xintercept=0),col='black')+
  geom_abline(intercept = alfa, slope = beta, col='dodgerblue3')+
  geom_vline(aes(xintercept=qelasticity),col='palevioletred4',linetype='dotted')+
  geom_hline(aes(yintercept=qfp(qelasticity)),col='palevioletred4',linetype='dotted')+
  geom_label(aes(x=qelasticity, y=qfp(qelasticity)), label=labl ,nudge_x = 25, nudge_y=15, size=5)+
  geom_label(aes(x=(qelasticity - 30) , y=(qfp(qelasticity)+30)), label='Elastic' ,nudge_x = 10, nudge_y=30, size=5)+
  geom_label(aes(x=(qelasticity + 30) , y= (qfp(qelasticity) - 30)), label='Not elastic' ,nudge_x =15, nudge_y=-30, size=5)+
  theme_bw()


```

### Maximize expanditure

The maximum expenditure is expected in the unit elasticity point, because that is the point, where 1% price increase will result 1% quantity loss, which means that the gain is approximately the same as the loss.


## TASK14: Flights to Budapest

* At price 40$ quantity is 300
* At price 60$ quantity is 280

### Demand equation

```{r demandeq, echo=FALSE}
p1<- 40
p2<- 60
q1<-300
q2<-280
beta <- (p2-p1) / (q2-q1)
alfa <- (-1)*q2*beta+p2
e1 <- (p1/q1) * (1/beta) 
e2 <- (p2/q2) * (1/beta)
lab1 <- paste(
  paste("(","):\n", sep=paste(as.character(q1),as.character(p1),sep=",")),
  paste("E1: ",substring(as.character(e1),1,6))
)
lab2 <- paste(
  paste("(","):\n", sep=paste(as.character(q2),as.character(p2),sep=",")),
  paste("E2: ",substring(as.character(e2),1,6))
)
##How to use equation dynmically?!
```
$$P=340-1*Q$$


```{r demandplot, echo=FALSE}
ggplot()+
  xlim(0, (-1*(alfa/beta) + 30))+
  ylim(-2,alfa+10)+
  labs(
    x='Q(antity)',
    y='P(rice)',
    title='Demand curve for Budapest flights'
  )+
  geom_hline(aes(yintercept=0),col='black')+
  geom_vline(aes(xintercept=0),col='black')+
  geom_abline(intercept = alfa, slope = beta, col='dodgerblue3')+
  geom_vline(aes(xintercept=q1),col='palevioletred4',linetype='dotted')+
  geom_hline(aes(yintercept=p1),col='palevioletred4',linetype='dotted')+
  geom_vline(aes(xintercept=q2),col='red4',linetype='dotted')+
  geom_hline(aes(yintercept=p2),col='red4',linetype='dotted')+
  geom_label(aes(x=(q2 - 20) , y=p2+20), label=lab2 ,nudge_x = -10, nudge_y=10, size=3)+
  geom_label(aes(x=q1 + 20  , y= p2 - 20), label=lab1 ,nudge_x = 10, nudge_y=10, size=3)+
  theme_bw()

```
