---
title: "Consumer and Firm Behavior, Assignment1"
author: "CSORNAI, Gyula"
date: "February 18, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment1 details 



###TASK9: Boris's morining coffee

#### Short summary

Blue line shows the maximum amount of coffe and milk combinations Boris can chose. 
At the maximum he can buy 18oz milk and 9oz coffe on 1$ coffee price: *MILK = 18-2COFFEE*.
However he prefers 4 part coffe, 1 part milk ( *MILK=1/4COFFEE*), so on this price his optimal quantities are

*1/4COFFE = 18-2 x COFFEE -> 2.25COFFEE = 18 ->* **COFFEE = 8, MILK=2**

If coffe prices goes from 1$ to 3.25 USD, his budget line changes:

*MILK=18-18/(9/3.25)COFFEE -> MILK=18-6.75COFFEE* 

so

*1/4COFFEE = 18-6.5COFFE -> 6.75COFFEE = 18 ->* **COFFEE=2.67, MILK=0.67**


#### Visualization


```{r pressure, echo=FALSE}
library(ggplot2)
icp2 <- (-1)*18/(9/3.25)
ggplot()+
  xlim(0,11)+
  ylim(-4,20)+
  labs(
    x='Price of coffee',
    y='Price of milk',
    title='Consumer choices of Boris related to coffee and milk'
  )+
  geom_hline(aes(yintercept=0),col='black')+
  geom_vline(aes(xintercept=0),col='black')+
  geom_abline(intercept = 18, slope = -2, col='dodgerblue3')+
  geom_abline(intercept = 18, slope = icp2, col='indianred')+
  geom_vline(aes(xintercept=8),col='chocolate4',linetype='dashed')+
  geom_hline(aes(yintercept=2),col='chocolate4',linetype='dashed')+
  geom_vline(aes(xintercept=2.67),col='palevioletred4',linetype='dotted')+
  geom_hline(aes(yintercept=0.67),col='palevioletred4',linetype='dotted')+
  geom_label(aes(x=0, y=18, label='Max milk volume\n for budget=18oz'),nudge_x = 1, nudge_y=1, size=3)+
  geom_label(aes(x=9, y=0, label='Max coffe volume\n for budget on\n 1$ price = 9oz'),nudge_x = 0, nudge_y=-2, size=3)+
  geom_label(aes(x=3, y=0, label='Max coffe volume\n for budget on\n 3.25$ price = 9oz'),nudge_x = 0, nudge_y=-2, size=3)+
  geom_label(aes(x=8, y=2, label='Preference point for budget\non 1$ coffee price: 8oz coffee,\n 2oz milk'),nudge_x = 1.5, nudge_y=2, size=3)+
  geom_label(aes(x=2.67, y=0.67, label='Preference point for budget\non 3.25$ coffee price:\n 2.67oz coffee, 0.67 oz milk '),nudge_x = 2, nudge_y=3, size=3)+
  theme_bw()
```

